---
title: 'Statistical Computing: Data Frames and Apply (TA VERSION)'
author: "TA: Daniel Zhu"
date: "Week of Tuesday September 12, 2023"
output: pdf_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(cache=TRUE, autodep=TRUE, cache.comments=TRUE)
```

**This week's agenda**: getting familiar with data frames; practicing how to use the apply family of functions.

States data set
===

Below we construct a data frame, of 50 states x 10 variables. The first 8 variables are numeric and the last 2 are factors. The numeric variables here come from the built-in `state.x77` matrix, which records various demographic factors on 50 US states, measured in the 1970s. You can learn more about this state data set by typing `?state.x77` into your R console. 

```{r}
state.df = data.frame(state.x77, Region=state.region, Division=state.division)
```

Q1. Basic data frame manipulations
===

- **1a.** Add a column to `state.df`, containing the state abbreviations that are stored in the built-in vector `state.abb`. Name this column `Abbr`. You can do this in (at least) two ways: by using a call to `data.frame()`, or by directly defining `state.df$Abbr`. Display the first 6 rows and all 11 columns of the new `state.df`.

```{r q1a}
state.df$Abbr=state.abb
head(state.df,6)
```

- **1b.** Remove the `Frost` column from `state.df`. You can do this in (at least) two ways: by using negative indexing, or by directly setting `state.df$Frost` to be `NULL`. Display the first 6 rows and all 10 columns of `state.df`.

```{r q1b}
state.df$Frost = NULL
#state.df=state.df[-7]
head(state.df,6)
```

- **1c.** Add two columns to `state.df`, containing the x and y coordinates (longitude and latitude, respectively) of the center of the states, that are stored in the (existing) list `state.center`. Hint: take a look at this list in the console, to see what its elements are named. Name these two columns `Center.x` and `Center.y`. Display the first 6 rows and all 12 columns of `state.df`.

```{r q1c}
state.df$Center.x=state.center$x
state.df$Center.y=state.center$y

head(state.df,6)
```

- **1d.** Make a new data frame which contains only those states whose longitude is less than -100. Do this in two different ways: using manual indexing, and `subset`(). Check that they are equal to each other, using an appropriate function call.

```{r q1d}
state.df.new1=subset(state.df, Center.x < -100)
state.df.new2=state.df[state.df$Center.x < -100,]

all(state.df.new1==state.df.new2)
```

- **1e.** Make a new data frame which contains only the states whose longitude is less than -100, and whose murder rate is above 9%. Print this new data frame to the console. Among the states in this new data frame, which has the highest average life expectancy? Determine this programmatically.

```{r q1e}
state.df.new3=state.df[state.df$Center.x < -100 & state.df$Murder>9,]
state.df.new3
```
**Answer: California**

Prostate cancer data set
===

Below we read in the prostate cancer data set that we looked in the last lab. You can remind yourself about what's been measured by looking back at the lab.

```{r}
pros.dat = 
  read.table("http://www.stat.cmu.edu/~ftownes/teaching/36350/F23/data/pros.dat")
```

Q2. Practice with the apply family
===

- **2a.** Using `sapply()`, calculate the mean of each variable. Also, calculate the standard deviation of each variable. Each should require just one line of code. Display your results.

```{r q2a}
apply(state.x77, MARGIN=2, FUN=mean)
apply(state.x77, MARGIN=2, FUN=sd)
```

- **2b.** Let's plot each variable against SVI. Using `lapply()`, plot each column, excluding SVI, on the y-axis with SVI on the x-axis. This should require just one line of code. **Challenge**: label the y-axes in your plots appropriately. Your solution should still consist of just one line of code and use an apply function. Hint: for this part, consider using `mapply()`.

```{r q2b}
lapply(pros.dat[-5], FUN=plot, x=pros.dat$svi)
```

- **2c.** Now, use `lapply()` to perform t-tests for each variable in the data set, between SVI and non-SVI groups. To be precise, you will perform a t-test for each variable excluding the SVI variable itself. For convenience, we've defined a function `t.test.by.ind()` below, which takes a numeric variable `x`, and then an indicator variable `ind` (of 0s and 1s) that defines the groups. Run this function on the columns of `pros.dat`, excluding the SVI column itself, and save the result as `tests`. What kind of data structure is `tests`? Print it to the console.

```{r}
t.test.by.ind = function(x, ind) {
  stopifnot(all(ind %in% c(0, 1)))
  return(t.test(x[ind == 0], x[ind == 1]))
}
```

```{r q2c}
tests=lapply(pros.dat[-5], FUN=t.test.by.ind, ind=pros.dat$svi)
```

- **2d.** Using `lapply()` again, extract the p-values from the `tests` object you created in the last question. Hint: first, take a look at the first element of `tests`, what kind of object is it, and how is the p-value stored? Second, write a custom function that takes as input the first element of `tests` and returns the p-value for just this one element. Now use what you've learned to extract p-values from the `tests` object.

```{r q2d}
tests[1]
class(tests)

lapply(tests, FUN=`[[`,"p.value")
```

Rio Olympics data set
===

Now we're going to examine data from the 2016 Summer Olympics in Rio de Janeiro, taken from https://github.com/flother/rio2016 (complete data on the 2020 Summer Olympics in Tokyo doesn't appear to be available yet). Below we read in the data and store it as `rio`.

```{r}
rio = read.csv("http://www.stat.cmu.edu/~ftownes/teaching/36350/F23/data/rio.csv")
```

Q3. More practice with data frames and apply
===

- **3a.** What kind of object is rio? What are its dimensions and columns names of `rio`? What does each row represent? Is there any missing data?

```{r q3a}
class(rio)
colnames(rio)

is.na(rio$height[1:50])
is.na(rio$weight[1:50])
```
**Answer: "rio" is a data frame with 12 cols and 11538 rows. Each row representing an athlete and their info at the Rio Olympics. Most of the entries in the "info" column are left empty. We also see that there are some values for height and weight of athletes but some of them are missing**

- **3b.** Use `rio` to answer the following questions. How many athletes competed in the 2016 Summer Olympics? How many countries were represented? What were these countries, and how many athletes competed for each one? Which country brought the most athletes, and how many was this? Hint: for a factor variable `f`, you can use `table(f)` see how many elements in `f` are in each level of the factor.

```{r q3b}
sum(table(rio$name))

nationalities<-unique(rio$nationality)
length(nationalities)

table(rio$nationality)[which.max(table(rio$nationality))]
```
**Answer: 11538 athletes total, from 207 unique countries, with the US bringing the most athletes with a total of 567**

- **3c.** How many medals of each type---gold, silver, bronze---were awarded at this Olympics? Are they equal? Is this result surprising, and can you explain what you are seeing?

```{r q3c}
sum(rio$gold[rio$gold>0])
sum(rio$silver[rio$silver>0])
sum(rio$bronze[rio$bronze>0])
```
**Answer: There were 666, 655, and 704 gold silver and bronze medals respectfully, it is a little surprise me because the totals are not equal but a possible explanation for this discrepancy is that some athletes would win their medals as teams, and there may be some missing data here to create that difference**

- **3d.** Create a column called `total` which adds the number of gold, silver, and bronze medals for each athlete, and add this column to `rio`. Which athlete had the most number of medals and how many was this? Gold medals? Silver medals? In the case of ties, here, display all the relevant athletes. 

```{r q3d}
total<-(rio$gold+rio$silver+rio$bronze)
rio=cbind(rio,total)
rio$name[which(rio$total==max(rio$total))]

rio$gold[which(rio$name=="Michael Phelps")]
rio$silver[which(rio$name=="Michael Phelps")]

rio[which.max(rio$gold),2]
rio[which.max(rio$gold),2]
max(rio$silver)
rio[(rio$silver==1),2]
```
**Answer: Michael Phelps had the most medals with 6 which 5 of were golds, 1 of which was silver. The max gold medals were 5 for MP and the max silvers were two, which was a total achieved many people**

- **3e.** Using `tapply()`, calculate the total medal count for each country. Save the result as `total.by.nat`, and print it to the console. Which country had the most number of medals, and how many was this? How many countries had zero medals?

```{r q3e}
total.by.nat=tapply(rio[,"total"], INDEX=rio$nationality, FUN=sum)
```
**Answer: USA had the most medals with 264, 120 countries had no medals**

- **3f.** Among the countries that had zero medals, which had the most athletes, and how many athletes was this? (Ouch!)
```{r q3f}
table(rio$nationality)[which(total.by.nat==0)]
max(table(rio$nationality)[which(total.by.nat==0)])
```
**Answer: Chile, with 42 athletes and also had no medals**

Q4. Young and old folks
===

- **4a.** The variable `date_of_birth` contains strings of the date of birth of each athlete. Use the `substr()` function to extract the year of birth for each athlete, and then create a new numeric variable called `age`, equal to 2016 - (the year of birth). (Here we're ignoring days and months for simplicity.) Hint: to extract the first 4 characters of a string `str`, you can use `substr(str, 1, 4)`. As always, you can also look at the help file for `substr()` for more details.

  Add the `age` variable to the `rio` data frame. variable  Who is the oldest athlete, and how old is he/she? Youngest athlete, and how old is he/she? In the case of ties, here, display all the relevant athletes. 

```{r q4a}
rio$age=2016-as.numeric(substr(rio$date_of_birth,1,4))

rio[which(rio$age==max(rio$age)),]
rio[which(rio$age==min(rio$age)),]
```
**Answer: there were ties for both the oldest and youngest athletes at the games, so the names have been displayed. Oldest athletes were 62 while youngest 14 yo which isane**

- **4b.** Answer the same questions as in the last part, but now only among athletes who won a medal.

```{r q4b}
winners=subset(rio,total>0)

winners[which(winners$age==max(winners$age)),]
winners[which(winners$age==min(winners$age)),]
```
**Answer: Youngest and oldest athletes to win medals were Nick Skelton at 59, and Qian Ren at 15**

- **4c.** Using a single call to `tapply()`, answer: how old are the youngest and oldest athletes, for each sport? 

```{r q4c}
tapply(rio$age,rio$sport,range)
```

- **4d.** You should see that your output from `tapply()` in the last part is a list, which is not particularly convenient. Convert this list into a matrix that has one row for each sport, and two columns that display the ages of the youngest and oldest athletes in that sport. The first 3 rows should look like this:
```
                  Youngest Oldest
aquatics                14     41
archery                 17     44
athletics               16     47
```
  You'll notice that we set the row names according to the sports, and we also set appropriate column names. Hint: `unlist()` will unravel all the values in a list; and `matrix()`, as you've seen before, can be used to create a matrix from a vector of values. After you've converted the results to a matrix, print it to the console (and make sure its first 3 rows match those displayed above).
  
```{r q4d}
not_list=unlist(tapply(rio$age,rio$sport,range))
range_matrix=matrix(not_list, ncol=2, byrow=T)
colnames(range_matrix)=c("Youngest","Oldest")
rownames(range_matrix)=as.character(unique(rio$sport))

range_matrix
```

Q5. Sport by sport
===

- **5a.** Create a new data frame called `sports`, which we'll populate with information about each sporting event at the Summer Olympics. Initially, define `sports` to contain a single variable called `sport` which contains the names of the sporting events in alphabetical order. Then, add a column called `n_participants` which contains the number of participants in each sport. Use one of the apply functions to determine the number of gold medals given out for each sport, and add this as a column called `n_gold`. Using your newly created `sports` data frame, calculate the ratio of the number of gold medals to participants for each sport. Which sport has the highest ratio? Which has the lowest?

```{r q5a}
sports<-data.frame(sport=rownames(table(rio$sport)),
                   n_participants=tapply(rio$id,INDEX=rio$sport,FUN=length), 
                   n_gold=tapply(rio$gold,INDEX=rio$sport,FUN=sum))

sports$n_gold/sports$n_participants
sports[which.max(sports$n_gold/sports$n_participants),]
sports[which.min(sports$n_gold/sports$n_participants),]
```
**Answer: Gymnastics has the highest winrate while golf is the lowest didnt know there was golf in the olympics**

- **5b.** Use one of the apply functions to compute the average weight of the participants in each sport, and add this as a column to `sports` called `ave_weight`. Important: there are missing weights in the data set coded as `NA`, but your column `ave_weight` should ignore these, i.e., it should be itself free of `NA` values. You will have to pass an additional argument to your apply call in order to achieve this. Hint: look at the help file for the `mean()` function; what argument can you set to ignore `NA` values? Once computed, display the average weights along with corresponding sport names, in decreasing order of average weight.

```{r q5b}
sports$avg_weight=tapply(rio$weight,rio$sport, FUN=mean, na.rm=TRUE)
sports[order(sports$avg_weight, decreasing=TRUE),c(1,4)]
```

- **5c.** As in the last part, compute the average weight of athletes in each sport, but now separately for men and women. You should therefore add two new columns, called `ave_weight_men` and `ave_weight_women`, to `sports`. Once computed, display the average weights along with corresponding sports, for men and women, each list sorted in decreasing order of average weight. Are the orderings roughly similar?

```{r q5c}
men=subset(rio,rio$sex=="male")
women=subset(rio,rio$sex=="female")

sports$avg_weight_men=tapply(men$weight,men$sport,FUN=mean,na.rm=TRUE)
sports$avg_weight_women=tapply(women$weight,women$sport,FUN=mean,na.rm=TRUE)

sports$avg_weight_men
sports$avg_weight_women
```
**Answer: I had trouble displaying the data but it seems like the weights are roughly the same but with the men as higher outliers while women were the lower values. For some indexes it seemed realities the same**
