---
title: 'Statistical Computing: Plotting Tools (TA VERSION)'
author: "TA: Daniel Zhu"
date: "Week of Tuesday October 31, 2023"
output: pdf_document
---

```{r, include=FALSE}
# A hook to wrap output based on a linewidth chunk option
# From https://github.com/yihui/knitr-examples/blob/master/077-wrap-output.Rmd
library(knitr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})
opts_chunk$set(cache=TRUE, autodep=TRUE, cache.comments=TRUE, linewidth=79)
```

 
**This week's agenda**: getting familiar with basic plotting tools; understanding the way layers work; recalling basic text manipulations; producing histograms and overlaid histograms; heatmaps.

Q1. Plot basics
===

- **1a.** Below is some code that is very similar to that from the lecture, but with one key difference. 

**Respond:** why does the `plot()` result with with `type="p"` look normal, but the `plot()` result with `type="l"` look abnormal, having crossing lines? 

Then modify the code below (hint: modify the definition of `x`) so that the lines on the second plot do not cross. 

```{r}
n = 50
set.seed(0)
x = runif(n, min=-2, max=2)
y = x^3 + rnorm(n)
plot(x, y, type="p")
plot(x, y, type="l")
```
**Answer: The "l" type plot appears distorted because the values lack proper sorting. R lacks the necessary information to connect the points in ascending order from -2 to 2; instead, it simply traces the data points as provided by the "runif" function, which generates random x values within the range of -2 to 2. This issue does not impact the "p" type plot, as it does not involve connecting points, making it look consistent regardless of the order in which it is drawn.**

```{r q1a}
n = 50
set.seed(0)
x = sort(runif(n, min=-2, max=2))
y = x^3 + rnorm(n)
plot(x, y, type="p")
plot(x, y, type="l")
```

- **1b.** The `cex` argument can used to shrink or expand the size of the points that are drawn. Its default value is 1 (no shrinking or expansion). Values between 0 and 1 will shrink points, and values larger than 1 will expand points. Plot `y` versus `x`, first with `cex` equal to 0.5 and then 3 (so, two separate plots). Give titles "Shrunken points", and "Expanded points", to the plots, respectively.

```{r q1b}
plot(x, y, cex = 0.5, main = "Shrunken points")
plot(x, y, cex = 3, main = "Expanded points")
```

- **1c.** The `xlim` and `ylim` arguments can be used to change the limits on the x-axis and y-axis, respectively. For example, each argument takes a vector of length 2, as in `xlim = c(-1, 0)`, to set the x limit to be from -1 to 0. 

Plot `y` versus `x`, set the x limit set to be from -1 to 1, and set the y limit set to be from -5 to 5. Assign x and y labels "Trimmed x" and "Trimmed y", respectively.

```{r q1c}
plot(x, y, xlim = c(-1,1), ylim = c(-5,5), 
     main = "Changing axis", 
     xlab = "Trimmed x", 
     ylab = "Trimmed y")
```

- **1d.** Plot `y` versus `x` so that you only show points whose x values are between -1 and 1. To do so, define two variables `x.trimmed` and `y.trimmed`, where `x.trimmed` is the subset of `x` between -1 and 1, and  `y.trimmed` is the corresponding subset of `y`. Then plot `y.trimmed` versus `x.trimmed` without setting `xlim` and `ylim`: now you should see that the y limit is (automatically) set as "tight" as possible. Hint: use logical indexing to define `x.trimmed`, `y.trimmed`.

```{r q1d}
x.trimmed = x[x < 1 & x > -1]
y.trimmed = x.trimmed^3 + rnorm(26)

plot(x.trimmed, y.trimmed)
```

- **1e.** The `pch` argument, recall, controls the point type in the display. In the lecture examples, we set it to a single number. But it can also be a vector of numbers, with one entry per point in the plot. So, e.g.,

```{r}
plot(1:25, 1:25, pch=1:25)
```
    
displays the first 25 point types. If `pch` is a vector whose length is shorter than the total number of points to be plotted, then its entries are recycled, as appropriate. Plot `y` versus `x`, with the point type alternating in between an empty square, filled square, plus symbol, and multiplication symbol (**note that the order of elements of the vector supplied to `pch` matters**).  

```{r q1e}
plot(x, y, pch = c(22, 15, 3, 4))
```

- **1f.** The `col` argument, recall, controls the color the points in the display. It operates similar to `pch`, in the sense that it can be a vector, and if the length of this vector is shorter than the total number of points, then it is recycled appropriately. Plot `y` versus `x`, and repeat the following pattern for the displayed points: a black empty circle, a blue filled circle, a black empty circle, a red filled circle.

```{r q1f}
plot(x, y, pch = c(1, 20), col = c("black","blue","black","red"))
```

Q2. Adding to plots
===

- **2a.** Produce a scatter plot of `y` versus `x`, and set the title and axes labels as you see fit. Then overlay on top a scatter plot of `y2` versus `x2`, using the `points()` function, where `x2` and `y2` are as defined below. In the call to `points()`, set the `pch` and `col` arguments appropriately so that the overlaid points are drawn as filled blue circles. 

```{r}
x2 = sort(runif(n, min = -2, max = 2))
y2 = x^2 + rnorm(n)
```

```{r q2a}
plot(x,y, ylim = c(-25,25), main = "Overlayed Scatterplots of Functions")
points(x2,y2, pch = 19, col = "blue")

legend(-1,25, legend = c("Equation 1: y = x^3 + rnorm(n)", 
                         "Equation 2: y = x^2 + rnorm(n)"),
           fill = c("white", "blue"))
```

- **2b.** Starting with your solution code from the last question, overlay a line plot of `y2` versus `x2` on top of the plot (which contains empty black circles of `y` versus `x`, and filled blue circles of `y2` versus `x2`), using the `lines()` function. In the call to `lines()`, set the `col` and `lwd` arguments so that the line is drawn in red, with twice the normal thickness. Look carefully at your resulting plot. 

**Respond**: Does the red line pass over the top of or underneath the blue filled circles? What do you conclude about the way R *layers* these additions to your plot?

```{r q2b}
plot(x,y, ylim = c(-25,25), main = "Overlayed Scatterplots of Functions")
points(x2,y2, pch = 19, col = "blue")

legend(-1,25, legend = c("Equation 1: y = x^3 + rnorm(n)", 
                         "Equation 2: y = x^2 + rnorm(n)"),
           fill = c("white", "blue"))
lines(x2,y2, col = "red", lwd = 2)
```
**Answer: The red line is drawn on top of the blue filled circles. From this R seems to follow a top-to-bottom layering approach in its code execution, where the code written last is positioned on top of the preceding code.**

- **2c.** Starting with your solution code from the last question, add a legend to the bottom right corner of the the plot using `legend()`. The legend should display the text: "Cubic" and "Quadratic", with corresponding symbols: an empty black circle and a filled blue circle, respectively. Hint: it will help to look at the documentation for `legend()`.

```{r q2c}
plot(x,y, ylim = c(-25,25), main = "Overlayed Scatterplots of Functions")
points(x2,y2, pch = 19, col = "blue")

legend(1,-5, legend = c("Cubic", "Quadratic"),
                fill = c("white", "blue"))
lines(x2,y2, col = "red", lwd = 2)
```

- **2d.** Produce a plot of `y` versus `x`, but with a gray rectangle displayed underneath the points, which runs has a lower left corner at `c(-2, qnorm(0.1))`, and an upper right corner at `c(2, qnorm(0.9))`. Hint: use `rect()` and consult its documentation. Also, remember how layers work; call `plot()`, with `type="n"` or `col="white"` in order to refrain from drawing any points in the first place, then call `rect()`, then call `points()`.

```{r q2d}
plot(x, y, col = "white")
rect(-2, qnorm(0.1), 2, qnorm(0.9), col = "lightblue")
points(x,y)
```

Fastest 100m sprint times
===

Below, we read in two data sets of the 1000 fastest times ever recorded for the 100m sprint, in men's and women's track., as seen in previous labs.

```{r}
sprint.m.df = read.table(
  file="https://www.stat.cmu.edu/~ftownes/teaching/36350/F23/data/sprint.m.txt", 
  sep="\t", quote="", header=TRUE)
sprint.w.df = read.table(
  file="https://www.stat.cmu.edu/~ftownes/teaching/36350/F23/data/sprint.w.txt", 
  sep="\t", quote="", header=TRUE)
```

Q3. Text manipulations, and layered plots
===

- **3a.** Define `sprint.m.times` to be the first 4 characters of the `Time` column of `sprint.m.df`, and  `sprint.m.years` to be the last 4 characters of the `Date` column of `sprint.m.df`. Hint: use `substr()`. Convert both to numeric vectors, and print the first 10 entries of each. 

```{r q3a}
sprint.m.times <- as.numeric(substr(sprint.m.df$Time, 1, 4))
sprint.m.years <- as.numeric(substr(sprint.m.df$Date, 
                                    nchar(sprint.m.df$Date) - 3, 
                                    nchar(sprint.m.df$Date)))
head(sprint.m.times, 10)
head(sprint.m.years, 10)
```

- **3b.** Plot `sprint.m.times` versus `sprint.m.years`. For the point type, use small, filled black circles. Label the x-axis "Year" and the y-axis "Time (seconds)". Title the plot "Fastest men's 100m sprint times". Using `abline()`, draw a dashed blue horizontal line at 9.95 seconds. Using `text()`, draw below this line, in text on the plot, the string "N men", replacing "N" here by the number of men who have run under 9.95 seconds. Your code should programmatically determine the correct number here, and use `paste()` to form the string. 

**Respond:** Comment on what you see visually, as per the sprint times across the years. What does the trend look like for the fastest time in any given year?

```{r q3b}
plot(sprint.m.years, sprint.m.times, pch = 20, 
     xlab = "Year", ylab = "Time (seconds)", 
     main = "Fastest men's 100m sprint times")
abline(a = 9.95, b = 0, col = "blue", lty = 2)

text(1995, 9.65, paste(as.character(length(sprint.m.times
                    [sprint.m.times < 9.95])), "men (under 9.95)"), 
                     col = "red", cex = 2)
```
**Answer: Sprint times have consistently decreased over the years, indicating that individuals are likely becoming faster with time. Notably, there is a declining trend in the fastest recorded time each year, with the peak performance occurring in 2009.**

- **3c.** Reproduce the previous plot, but this time, draw a light blue rectangle underneath all of the points below the 9.95 second mark. The rectangle should span the entire region of the plot below the horizontal line at $y=9.95$. And not only the points of sprint times, but the blue dashed line, and the text "N men" (with "N" replaced by the appropriate number) should appear *on top* of the rectangle. Hint: use `rect()` and layering as appropriate.

```{r q3c}
plot(sprint.m.years, sprint.m.times, col = "white", 
     xlab = "Year", ylab = "Time (seconds)", 
     main = "Fastest men's 100m sprint times")
abline(a = 9.95, b = 0, col = "blue", lty = 2, lwd = 2)

rect(1965, 9.5, 2025, 9.95, col = "lightblue")
points(sprint.m.years, sprint.m.times, pch = 20)


text(1995, 9.98, paste(as.character(length(sprint.m.times
                    [sprint.m.times < 9.95])), "men"), 
                     col = "red", cex = 2)
```
**Confused why you would want to put the printout somewhere you can't see but thems the rules**

- **3d.** Repeat Q3a but for the women's sprint data, with one critical difference---to form the vector of times, take the **first 5 characters** of the `Time` column---and arrive at vectors `sprint.w.times` and `sprint.w.years`. Then repeat Q3c for this data, but with the 9.95 second cutoff being replaced by 10.95 seconds, the rectangle colored pink, and the dashed line colored red. 

**Respond**: Comment on the differences between this plot for the women and your plot for the men, from Q4c. In particular, is there any apparent difference in the trend for the fastest sprint time in any given year?

```{r q3d}
sprint.w.times <- as.numeric(substr(sprint.w.df$Time, 1, 5))
sprint.w.years <- as.numeric(substr(sprint.w.df$Date, 
                                    nchar(sprint.w.df$Date) - 3, 
                                    nchar(sprint.w.df$Date)))
plot(sprint.w.years, sprint.w.times, col = "white", 
     xlab = "Year", ylab = "Time (seconds)", 
     main = "Fastest women's 100m sprint times")
abline(a = 10.95, b = 0, col = "red", lty = 2, lwd = 2)

rect(1970, 10, 2025, 10.95, col = "pink")
points(sprint.w.years, sprint.w.times, pch = 20)

text(1998, 10.55, paste(as.character(length(sprint.w.times
                    [sprint.w.times < 10.95])), "women"), col = "blue", cex = 2)
```
**Answer: The plot appears to have a broader distribution, indicating a greater amount of data for earlier years compared to the men's data frame. It's worth noting that the negative slope between time and year is less steep than the slope for the men's data. The trend in the fastest sprint times, year by year, displays a descending pattern, albeit with a less pronounced rate of decline compared to the men's fastest sprint times.**

Q4. More text manipulations, and histograms
===

- **4a.** Extract the birth years of the sprinters from the data frame `sprint.m.df`. To do so, define a character vector `sprint.m.byears` to contain the last 2 characters of the `Birthdate` column of `sprint.m.df`. Then convert `sprint.m.byears` into a numeric vector, add 1900 to each entry, and redefine `sprint.m.byears` to be the result. Repeat the same, but for the women's data, arriving at a vector called `sprint.w.byears`. Print the first 10 entries of each vector.

```{r q4a}
sprint.m.byears <- as.numeric(substr(sprint.m.df$Birthdate, 
                              nchar(sprint.m.df$Birthdate) - 1, 
                              nchar(sprint.m.df$Birthdate)))
sprint.m.byears <- 1900 + sprint.m.byears

sprint.w.byears <- as.numeric(substr(sprint.w.df$Birthdate, 
                              nchar(sprint.w.df$Birthdate) - 1, 
                              nchar(sprint.w.df$Birthdate)))
sprint.w.byears <- 1900 + sprint.w.byears

sprint.m.byears[1:10]
sprint.w.byears[1:10]
```

- **4b.** What are the ranges of birth years you computed in the last part, for the men's and women's data? Do you see any problems with this? You should! Fix the errant entries in `sprint.m.byears` and `sprint.w.byears` using simple indexing and arithmetic. Print the first 10 entries of each vector.

Hint: none of these athletes were born before 1921.

```{r q4b}
range(sprint.m.byears)
range(sprint.w.byears)

sprint.m.byears[which(sprint.m.byears >= 1900 & sprint.m.byears < 1921)] =  sprint.m.byears[which(sprint.m.byears >= 1900 & sprint.m.byears < 1921)] + 100
sprint.w.byears[which(sprint.w.byears >= 1900 & sprint.w.byears < 1921)] = sprint.w.byears[which(sprint.w.byears >= 1900 & sprint.w.byears < 1921)] + 100
```
**Answer: Range was 1900 to 1999, the problem is the range as the minimum value is 1900 which is not possible**

- **4c.** Now that you have fixed their birthdates, create a vector `sprint.m.ages` containing the age (in years) of each male sprinter when their sprint time was recorded. Do the same for the female sprinters, resulting in `sprint.w.ages`. Print the first 10 entries of each vector.

Hint: use `sprint.m.years` and `sprint.w.years`.

```{r q4c}
sprint.m.ages <- sprint.m.years - sprint.m.byears
sprint.w.ages <- sprint.w.years - sprint.w.byears

sprint.m.byears[1:10]
sprint.w.byears[1:10]
```

- **4d.** Using one of the apply functions, compute the average sprint time for each age in `sprint.m.ages`, calling the result `time.m.avg.by.age`. Similarly, compute the analogous quantity for the women, calling the result `time.w.avg.by.age`. Are there any ages for which the men's average time is faster than 10.0 seconds, and if so, which ones? Are there any ages for which the women's average time is faster than 10.9 seconds, and if so, which ones?

```{r q4d}
sprint.m.df$Age <- sprint.m.ages
sprint.w.df$Age <- sprint.w.ages

time.m.avg.by.age <- tapply(as.numeric(sprint.m.df$Time), 
                            sprint.m.df$Age, FUN = mean, na.rm = TRUE)
time.w.avg.by.age <- tapply(as.numeric(sprint.w.df$Time), 
                            sprint.w.df$Age, FUN = mean, na.rm = TRUE)

time.m.avg.by.age[which(time.m.avg.by.age < 10.0)]
time.w.avg.by.age[which(time.w.avg.by.age < 10.9)]
```
**Answer: Print out is a little hard to read but it seems like men age 41 had a faster average while from women it was ages 29, 33, 35, and 37**

- **4e.** Plot a histogram of `sprint.m.ages`, with break locations occuring at every age in between 17 and 40. Color the histogram to your liking; label the x-axis, and title the histogram appropriately. 

**Respond**: What is the mode, i.e., the most common age? Also, describe what you see around the mode: do we see more sprinters who are younger, or older? 

```{r q4e}
#validSprint.m.ages <- sprint.m.ages[!is.na(sprint.m.ages) & 
#                       sprint.m.ages >= 17 & sprint.m.ages <= 40]


#hist(validSprint.m.ages, breaks = 17:40, col = "skyblue", xlab = "Age", 
#     main = "Sprinter Ages")

#ageFreq <- table(validSprint.m.ages)
#modeAge <- as.numeric(names(ageFreq[which.max(ageFreq)]))
```
**Answer: umm, ask Professor why the code is being funny**

Q5. Maungawhau volcano and heatmaps (optional)
===

- **Challenge.** The `volcano` object in R is a matrix of dimension 87 x 61. It is a digitized version of a topographic map of the Maungawhau volcano in Auckland, New Zealand. Plot a heatmap of the volcano using `image()`, with 25 colors from the terrain color palette. 

```{r q5chal1}
image(volcano, col = terrain.colors(25), main = "Heatmap of Maungawhau volcano")
```

- **Challenge.** Each row of `volcano` corresponds to a grid line running east to west. Each column of `volcano` corresponds to a grid line running south to north. Define a matrix `volcano.rev` by reversing the order of the rows, as well as the order of the columns, of `volcano`. Therefore, each row `volcano.rev` should now correspond to a grid line running west to east, and each column of `volcano.rev` a grid line running north to south. 

```{r q5chal2}
volcano.rev <- volcano[87:1, 61:1]
```

- **Challenge.** If we printed out the matrix `volcano.rev` to the console, then the elements would follow proper geographic order: left to right means west to east, and top to bottom means north to south. Now, produce a heatmap of the volcano that follows the same geographic order. Hint: recall that the `image()` function rotates a matrix 90 degrees counterclockwise before displaying it; and recall the function `clockwise90()` from the lecture, which you can copy and paste into your code here. Label the x-axis "West --> East", and the y-axis "South --> North". Title the plot "Heatmap of Maungawhau volcano".

```{r q5chal3}
image(volcano.rev, col = terrain.colors(25), 
      main = "Heatmap of Maungawhau volcano", 
      xlab = "West to East", ylab = "South to North")
```

- **Challenge.** Reproduce the previous plot, and now draw contour lines on top of the heatmap.

```{r q5chal4}
image(volcano.rev, col = terrain.colors(25), 
      main = "Heatmap of Maungawhau volcano", 
      xlab = "West to East", ylab = "South to North")
contour(volcano.rev, add = TRUE)
```

- **Challenge.** The function `filled.contour()` provides an alternative way to create a heatmap with contour lines on top. It uses the same orientation as `image()` when plotting a matrix. Use `filled.contour()` to plot a heatmap of the volcano, with (light) contour lines automatically included. Make sure the orientation of the plot matches proper geographic orientation, as in the previous question. Use a color scale of your choosing, and label the axes and title the plot appropriately. It will help to consult the documentation for `filled.contour()`.

```{r q5chal5}
filled.contour(volcano.rev, color.palette = terrain.colors, 
               main = "Heatmap of Maungawhau volcano", 
               xlab = "West to East", ylab = "South to North")
```
